---
title: "Covid91 vaccine study Final2024"
author: "Larissa Vera!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##Methods

$H_0$: The vaccine is not effective in preventing Covid 921 in males

$H_A$: The vaccine effective in preventing Covid 921 in males.

The helper file used is cat~ cat.

## Males

## Graphical results

```{r}
Male <- FinalData[FinalData$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

The barchart shows that males who recieved the vaccine are less likely to contact covid 921

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

the numerical data shows that within the placebo group 73.80% of males are infected with covid 921. While the males who received the vaccine only 26.20% contracted covid 921. Therefore the males who received the vaccine reduced the risk of contracting covid 921.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Females

## Graphical results

```{r}
Female <-FinalDat[FinalDat$sex=="F",]
```


```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```


### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

## LGBTQ

## Graphical results

```{r}
barchartGC(~LGBTQ + treatment,data=FinalData)
barchartGC(~LGBTQ + treatment,data=FinalData, type="percent")

```



### Numerical Summaries  

```{r}
table1 <- xtabs(~LGBTQ + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```



## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

## Druggies

## Graphical results
```{r}
barchartGC(~DrugUser + treatment,data=FinalData)
barchartGC(~DrugUser + treatment,data=FinalData, type="percent")

```

### Numerical Summaries  

```{r}
table1 <- xtabs(~DrugUser + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```


## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

# Overall Results and Conclusions

