---
title: "Covid91 vaccine study Final2024"
author: "Larissa Vera!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##Methods

$H_0$: The vaccine is not effective in preventing Covid 921 in males

$H_A$: The vaccine effective in preventing Covid 921 in males.

The helper file used is cat~ cat.

## Males

## Graphical results

```{r}
Male <- FinalData[FinalData$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

The barchart shows that males who recieved the vaccine are less likely to contact covid 921

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

the numerical data shows that within the placebo group 73.80% of males are infected with covid 921. While the males who received the vaccine only 26.20% contracted covid 921. Therefore the males who received the vaccine reduced the risk of contracting covid 921.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Square p value is 0.00000000000000022. Fisher exact test shows the p value is 0.00000000000000022 which is lower than 0.05 so we reject the null hypothesis.

## Conclusion

The bar chart shows you are less likely to contract covid 921 if you received the vaccine. The numerical data shows within the placebo 73.80% of males were infected while the males who received the vaccine only 26.20% contracted covid 921. Therefore this indicates that those how received the vaccine significantly reduced the risk of contracting covid 921. The fisher exact test shows the p value is 0.00000000000000022 which is lower than 0.05 so we reject the null hypothesis.

## Females

how effective is the vaccine on females?

## Method

$H_0$: The vaccine is not effective in preventing covid 921 in females

$H_A$: The vaccine is effective in preventing covid 921 in females

The helper file used is cat~cat

## Graphical results

```{r}
Female <-FinalDat[FinalDat$sex=="F",]
```


```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```

The bar chart shows females that received the vaccines are less likely to contract covid 921

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows within the placebo group 64.93% of females were infected with covid 921. While the females who received the vaccine only 35.07% contracted covid 921. Therefore this indicates that the females who received the vaccine significantly reduced the risk of contracting covid 921.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Square test shows that the p value is 0.00000001914 and the Fisher exact test shows that the p value is 0.00000001246 which is lower than 0.05 so we reject the null hypothesis.

## Conclusion

The bar chart shows females that received the vaccine are less likely to contract covid 921. The numerical data shows within the placebo group 64.93% of females got infected with covid 921. While the females who received the vaccine only 35.07% contracted covid 921. Therefore the females who received the vaccine significantly reduced the risk of contracting covid 921. The Fisher exact test shows that the p value is 0.00000001912 which is lower than 0.05 so we reject the null hypothesis. 

## LGBTQ

How effective is the vaccine on LGBTQ?

## Method 

$H_0$: The vaccine is not effective in preventing covid 921 in LGBTQ

$H_A$: The vaccine is effective in preventing covid 921 in LGBTQ

The helper file used is cat~cat

## Graphical results

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ=='gay',]
```


```{r}
barchartGC(~infected + treatment,data= LGBTQ)
barchartGC(~infected + treatment,data= LGBTQ, type="percent")

```

The bar chart shows that the LGBTQ community are less likely to contract covid 921 if you received the vaccine. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 80% of the LGBTQ community got infected with covid 921. While the LGBTQ community that received the vaccine only 20% contracted covid 921.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Druggies

## Graphical results
```{r}
barchartGC(~DrugUser + treatment,data=FinalData)
barchartGC(~DrugUser + treatment,data=FinalData, type="percent")

```

### Numerical Summaries  

```{r}
table1 <- xtabs(~DrugUser + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```


## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

# Overall Results and Conclusions

